---
title: "Spring/Summer Chinook parr survival estimates"
author:

- Michelle A. Briggs:
    email: michelleb@nezperce.org
    institute: npt_mccall
    correspondence: yes
- Ryan N. Kinzer:
    email: ryank@nezperce.org
    institute: npt_mccall
    correspondence: yes
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    theme: cerulean
    toc: no
    toc_depth: 4
    toc_float:
      collapsed: no
      smooth_scroll: yes
    number_sections: yes
    #css: "../templates/styles.css"
    pandoc_args:
     - "--lua-filter=templates/scholarly-metadata.lua"
     - "--lua-filter=templates/author-info-blocks.lua"
     - "--lua-filter=templates/pagebreak.lua"
  # bookdown::word_document2:
  #   toc: yes
  #   toc_depth: 4
  #   reference_docx: "../templates/Word_template_draft.docx"
institute:
- npt_mccall: Nez Perce Tribe, Department of Fisheries Resources Management, 14054 Burr Dr., McCall, Idaho, 83638, USA
csl: "templates/american-fisheries-society.csl"
always_allow_html: yes
editor_options: 
  chunk_output_type: console
bibliography: references.bib
---

```{r, include=FALSE}
library(tidyverse)
library(kableExtra)

plot_theme <- readRDS("templates/plot_theme_v2.RDS")
```

```{r, include = FALSE}
ch <- readRDS("data/parr_tagging/ch_parr_2025.RDS")
estimates <- readRDS("data/parr_tagging/parr_estimates_2025.RDS")
obs_tbl <- readRDS("data/parr_tagging_obs_tbl.RDS")

```

### Methods

Survival was estimated for juvenile spring/summer (sp/sum) Chinook salmon tagged in 2025. Chinook salmon parr were tagged with PIT tags in July and August of 2025 at fifteen locations in the Snake River Basin. The number of fish tagged by location varied from 15 to 1,499 ( Table \@ref(tab:tab-1)). Fish were subsequently detected at in-stream PIT tag arrays, at Lower Granite Dam, and at sites downstream of Lower Granite Dam in the Snake and Columbia rivers. Detections occurred at ten in-stream PIT tag arrays, with each array detecting between 1 and 1,055 tagged juvenile sp/sum Chinook salmon (Table \@ref(tab:tab-2)). Detections at Lower Granite Dam (LGR) included detections in the juvenile bypass system and the spillway. Detections at sites downstream of LGR were pooled. Detections occurred from July 2025 to August 2026 (Figure \@ref(fig:emigration-plot)).

Reach-specific survival (defined as the probability of survival from one detection site to the subsequent downstream detection site) was estimated using the M-array formulation of the Cormack-Jolly-Seber (CJS) model [@burnham1987]. The M-array formulation of the CJS model estimates survival for each reach from the release site to Lower Granite Dam and detection probability for each in-stream array and for Lower Granite Dam. For the final reach (i.e., downstream of Lower Granite Dam) a single parameter $\lambda$ , representing the product of final survival and detection, is estimated, because survival and detection cannot be estimated independently. Survival was not estimated for fish tagged in Bear Creek (BEAR6C) due to low sample size.

```{r, include = FALSE}
#tables of number of fish by release site and detection sites

LGR <- c('GRJ', 'GRS') #s = spillway, #j = juvenile bypass
Down <- c("GOJ", "LMJ", "ICH", "MCJ", "JDJ", "B2J", "BCC",
          "PD5", "PD6", "PD7", "PD8", "PDO", "PDW", "TWX")
obs_sites <- unique(estimates$estimates$from_site)
release_sites <- unique(ch$release_site)

site_names <- c("Bear Creek", "Hurricane Creek", 
                "Wallowa River", "Capehorn Creek", 
                "Chamberlin Creek", "Marsh Creek", 
                "Burnt Log Creek", "Grouse Creek", 
                "Johnson Creek", "Lake Creek", 
                "East Fork South Fork Salmon River", 
                "South Fork Salmon River", "Secesh River", 
                "Summit Creek", "Lolo Creek")

#group release sites by area
gri <- c("BEAR6C", "HURRIC", "WALLOR") 
sf_sal <- c("BURNLC", "GROUSC", "JOHNSC", "LAKEC", "SAEFSF", "SALRSF", "SECESR", "SUMITC")
mf_sal <- c("CHAMBC", "MARSHC", "CAPEHC")
clear <- c("LOLOC")

n_rel_site <- ch %>%
  select(tag_code, release_site) %>%
  distinct() %>%
  group_by(release_site) %>%
  count() %>%
  mutate(mpg = case_when(
    `release_site` %in% gri  ~ "Grande Ronde/Imnaha",
    `release_site` %in% sf_sal   ~ "SF Salmon",
    `release_site` %in% mf_sal ~ "MF Salmon",
    `release_site` %in% clear    ~ "Wet Clearwater")) %>%
  rename('Release site' = release_site, 
         'Number of tagged fish' = n) %>%
  arrange(mpg) 

n_rel_site <- cbind(n_rel_site, site_names)
n_rel_site <- n_rel_site %>%
  rename('Site name' = ...4)

site_by_mpg <- n_rel_site %>%
  select('Release site', mpg)

grp_site <- rle(as.character(n_rel_site$mpg))
index_site <- setNames(grp_site$lengths, grp_site$values)
n_rel_site_tbl <- n_rel_site %>%
  select(-mpg)

detect_site_key <- obs_tbl %>%
  mutate(obs_name = map(obs_loc, names)) %>%
  select(release_site, obs_name) %>%
  unnest(obs_name) %>%
  group_by(obs_name) %>%
  summarise(release_sites = paste(unique(release_site), collapse = ", "),  .groups = "drop") %>%
  filter(obs_name %in% obs_sites) %>%
  mutate(release_sites = if_else(obs_name == "LGR", "All", 
                                 if_else(obs_name == "Down", "All", release_sites))) %>%
  rename('Detection site' = obs_name)

n_detect_site <- ch %>%
  mutate(detection_site = if_else(obs_site %in% LGR, "LGR", 
                                         if_else(obs_site %in% Down, "Down", obs_site))) %>%
  select(tag_code, detection_site) %>%
  distinct() %>%
  group_by(detection_site) %>% 
  filter(detection_site %in% obs_sites) %>%
  filter(!(detection_site %in% release_sites)) %>%
  count() %>%
  rename('Detection site' = detection_site, 
         'Number of detected fish' = n) %>%
  left_join(detect_site_key, by = 'Detection site') %>%
  rename('Associated release sites' = release_sites)


```

```{r, include = F}
#tables of survival and detection estimates by release site

survival <- estimates$estimates %>%
  filter(metric == "survival", to_site != "Down")

detection <- estimates$estimates %>%
  filter(metric == "detection") %>%
  filter(!(is.na(marray_est)))

cumulative <- estimates$estimates %>%
  filter(metric == "cumulative") %>%
  filter(!is.na(marray_est))

survival_tbl <- survival %>%
  select(release_site, interval, from_site, to_site, marray_est, marray_lcl, marray_ucl) %>%
  mutate(marray_est = round(marray_est, 3),
         marray_lcl = round(marray_lcl, 3), 
         marray_ucl = round(marray_ucl, 3)) %>%
  unite("CI", marray_lcl:marray_ucl, sep = " - ") %>%
  rename('Release site' = release_site, 
         Interval = interval, 
         'From site' = from_site, 
         'To site' = to_site, 
         Estimate = marray_est, 
         '95% CI' = CI
         ) %>%
  left_join(site_by_mpg, by = "Release site") %>%
  arrange(mpg)

grp_sur <- rle(as.character(survival_tbl$'Release site'))
index_sur <- setNames(grp_sur$lengths, grp_sur$values)

mpg_rle   <- rle(survival_tbl$mpg)
index_mpg <- setNames(mpg_rle$lengths, mpg_rle$values)

survival_tbl <- survival_tbl %>% select(-'Release site')

detection_tbl <- detection %>%
  select(release_site, interval, from_site, marray_est, marray_lcl, marray_ucl) %>%
  mutate(marray_est = round(marray_est, 3),
         marray_lcl = round(marray_lcl, 3), 
         marray_ucl = round(marray_ucl, 3)) %>%
  unite("CI", marray_lcl:marray_ucl, sep = " - ") %>%
  rename('Release site' = release_site, 
         Interval = interval, 
         'Site' = from_site, 
         Estimate = marray_est, 
         '95% CI' = CI
         ) %>%
  left_join(site_by_mpg, by = "Release site") %>%
  arrange(mpg)

grp_det <- rle(as.character(detection_tbl$'Release site'))
index_det <- setNames(grp_det$lengths, grp_det$values)

detection_tbl <- detection_tbl %>% select(-'Release site')

cumulative_tbl <- cumulative %>%
  select(release_site, interval, to_site, marray_est, marray_lcl, marray_ucl) %>%
  mutate(marray_est = round(marray_est, 3),
         marray_lcl = round(marray_lcl, 3), 
         marray_ucl = round(marray_ucl, 3)) %>%
  unite("CI", marray_lcl:marray_ucl, sep = " - ") %>%
  rename('Release site' = release_site, 
         Interval = interval, 
         'To site' = to_site, 
         Estimate = marray_est, 
         '95% CI' = CI
         ) %>%
  left_join(site_by_mpg, by = "Release site") %>%
  arrange(mpg)

grp_cumulative <- rle(as.character(cumulative_tbl$'Release site'))
index_cumulative <- setNames(grp_cumulative$lengths, grp_cumulative$values)

cumulative_tbl <- cumulative_tbl %>% select(-'Release site')

```

```{r, include=F}
#emigration timing figures

detection_timing <- ch %>%
  filter(event_type != "Mark") %>% #filter out marking events
  mutate(detection_site = if_else(obs_site %in% LGR, "LGR", 
                                         if_else(obs_site %in% Down, "Down", obs_site))) %>%
  filter(detection_site %in% obs_sites) %>%
  filter(!(detection_site %in% release_sites)) %>%
  select(tag_code, detection_site, first_obs, release_site, mark_release_date) %>%
  arrange(first_obs)

detection_timing <- detection_timing %>%
  mutate(fish_num = 1) %>%
  group_by(release_site, detection_site) %>%
  mutate(cumul = cumsum(fish_num), 
         cum_prop = cumsum(fish_num)/sum(fish_num))

emigration_timing_plot <- detection_timing %>%
  filter(release_site != "BEAR6C") %>%
  filter(!(detection_site %in% c("CWR", "Down"))) %>%
  filter(!(detection_site == "ZEN" & release_site == "JOHNSC")) %>% #filter out 1 JC fish that was detected at ZEN
  ggplot(aes(x = first_obs, y = cum_prop)) + 
  geom_line(aes(color = release_site), linewidth = 1) +
  #scale_color_viridis_d(option = "viridis", name = "Release site") +
  facet_wrap(~detection_site, ncol = 3) + 
  plot_theme + 
  theme(strip.background = element_rect(fill = "grey", color = "black"), 
          axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ylab("Cumulative proportion") + 
  xlab("Date") + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) + 
  scale_x_date(expand = c(0, 0)) + 
  labs(color = "Release site")

```

```{r tab-1, echo = F, warning = F}
#knitr::kable(n_rel_site) #%>%
  #kable_styling(full_width = TRUE)

kableExtra::kbl(n_rel_site_tbl, 
                caption = "The number of juvenile sp/sum Chinook salmon tagged by release site in 2025. Release sites are grouped by major population group (MPG)." ) %>%
  #select(-area) %>%
  kable_styling(full_width = TRUE) %>%
  pack_rows(index = index_site)
```

```{r tab-2, echo = F, warning = F}
knitr::kable(n_detect_site, 
             caption = "The number of juvenile sp/sum Chinook salmon detected at instream detection sites during downstream migration. Detections at Down represent detections at all sites downstream of Lower Granite Dam combined. Fish tagged and released at the associated release sites will pass the indicated detection site during downstream migration.") #%>%
 # kable_styling(full_width = TRUE)
```

```{r emigration-plot, echo = F, fig.width=12, fig.height=14, fig.cap="Cumulative emigration timing of juvenile sp/sum Chinook salmon during downstream migration at each in-stream array and at Lower Granite Dam. Each color represents a unique tag group. Emigration timing is not shown for CWR, because only one fish was detected."}
emigration_timing_plot
```

### Results and discussion

Reach-specific estimates of survival for sp/sum Chinook parr were variable, and many estimates had broad confidence intervals (i.e., 95% CI 0 - 1, Figure \@ref(fig:surv-plot), Table \@ref(tab:tab-3)). Estimates of detection probability varied by detection site and by release group, with a minimum estimated detection probability of 0.002 for WR1 for the Hurricane Creek release group and a maximum detection probability of 0.935 for MAR for the Marsh Creek release group. Estimates of cumulative survival to Lower Granite Dam varied from 0.036 (95% CI 0.016 - 0.978) at Grouse Creek to 0.36 (95% CI 0.10 - 1.0) at Marsh Creek (Figure \@ref(fig:cumulative-plot), Table \@ref(tab:tab-5)). The average cumulative survival to Lower Granite Dam was 0.19, which is similar to estimates reported in previous research [@lamb2026].

Release groups with low sample sizes tended to have lower precision of reach-specific survival estimates. For example, Burnt Log Creek, Grouse Creek, and Summit Creek had sample sizes of 52, 168, and 174 tagged fish, respectively; each of these release groups had imprecise survival estimates. Estimates of detection probability for release groups with lower sample sizes also showed lower precision (Figure \@ref(fig:det-plot), Table \@ref(tab:tab-4)). The target sample size was a minimum of 300 tagged fish per release group, and the results suggest that reaching the target sample size may improve survival estimates. In addition to low sample sizes, low detection probabilities at some in-stream arrays, such as SFG, may be reducing the precision associated with reach-specific survival estimates.

To improve survival estimates, future modeling efforts will leverage shared detection sites among release groups by estimating a single detection probability per detection site. This approach will also reduce variability in estimates of detection probability for a single detection site. For example, estimates of detection probability for ZEN varied from 0.135 to 0.670 for different release groups (Table \@ref(tab:tab-4)), although emigration timing appears to be similar among groups (Figure \@ref(fig:emigration-plot)). Using multiple release groups to estimate a single detection probability may improve the precision of model estimates by increasing the sample size associated with downstream detection sites.

```{r tab-3, echo = F, warning = F}
survival_tbl %>%
  select(-mpg) %>%
  kbl(align = "c", 
      caption = "Reach-specific estimates of survival for juvenile sp/sum Chinook salmon during downstream migration to Lower Granite Dam.") %>%
  kable_styling(full_width = TRUE) %>%
  pack_rows(
    index = index_mpg,                                   # outer: area headings
    label_row_css = "text-align:center; font-weight:bold; background-color:#e9e9e9;",
    latex_align = "c"                                     # for PDF output
  ) %>%
  pack_rows(index = index_sur)                        # inner: release site groups


```

```{r tab-4, echo = F, warning = F}
detection_tbl %>%
  select(-mpg) %>%
  kbl(align = "c", 
      caption = "Estimates of detection probability for juvenile sp/sum Chinook salmon at in-stream detection sites during downstream migration to Lower Granite Dam. Detection probability was estimated independently for each release group.") %>%
  kable_styling(full_width = TRUE) %>%
  pack_rows(
    index = index_mpg,                                   # outer: area headings
    label_row_css = "text-align:center; font-weight:bold; background-color:#e9e9e9;",
  ) %>%
  pack_rows(index = index_det) 
```

```{r tab-5, echo = F, warning = F}
cumulative_tbl %>%
  select(-mpg) %>%
  kbl(align = "c", 
      caption = "Estimates of cumulative survival probability for juvenile sp/sum Chinook salmon during downstream migration to Lower Granite Dam.") %>%
  kable_styling(full_width = TRUE) %>%
  pack_rows(
    index = index_mpg,                                   # outer: area headings
    label_row_css = "text-align:center; font-weight:bold; background-color:#e9e9e9;",
  ) %>%
  pack_rows(index = index_cumulative) 
```

```{r, include = F}
#survival figures

surv_plot <- survival %>%
    #filter(release_site %in% site_subset) %>%
    unite(col = "int_label", from_site, to_site, sep = " to ", remove = FALSE) %>%
    mutate(int_label = fct_reorder(int_label, interval)) %>%
    left_join(site_by_mpg %>% rename(release_site = 'Release site'), 
              by = "release_site") %>%
    arrange(mpg) %>%
    #mutate(release_site = fct_reorder(release_site))
    ggplot(aes(x = int_label, y = marray_est)) + 
    geom_errorbar(aes(ymin = marray_lcl, ymax = marray_ucl, color = mpg), width = 0.1, linewidth = 1) +
    geom_point(size = 4, aes(fill = mpg), shape = 21) + 
    plot_theme + 
    facet_wrap(~fct_inorder(release_site), scales = "free_x", ncol = 3) +
    scale_y_continuous(limits = c(0, 1)) + 
    scale_color_viridis_d(option = "viridis", name = "MPG") +
    scale_fill_viridis_d(option = "viridis", name = "MPG") +
    theme(strip.background = element_rect(fill = "grey", color = "black"), 
          axis.text.x = element_text(angle = 45, hjust = 1)) + 
    ylab("Survival") + 
    xlab(" ")

#detection figures
detection_plot <- detection %>%
    mutate(site = fct_reorder(from_site, interval)) %>%
    left_join(site_by_mpg %>% rename(release_site = 'Release site'), 
              by = "release_site") %>%
    arrange(mpg) %>%
    ggplot(aes(x = site, y = marray_est)) + 
    geom_errorbar(aes(ymin = marray_lcl, ymax = marray_ucl, color = mpg), width = 0.1, linewidth = 1) +
    geom_point(size = 4, aes(fill = mpg), shape = 21) + 
    plot_theme + 
    facet_wrap(~fct_inorder(release_site), scales = "free_x", ncol = 3) +
    scale_y_continuous(limits = c(0, 1)) + 
    scale_color_viridis_d(option = "viridis", name = "MPG") +
    scale_fill_viridis_d(option = "viridis", name = "MPG") +
    theme(strip.background = element_rect(fill = "grey", color = "black"), 
          axis.text.x = element_text(angle = 45, hjust = 1)) + 
    ylab("Detection probability") + 
    xlab("Site")

cumulative_plot <- cumulative %>%
    mutate(site = fct_reorder(to_site, interval)) %>%
    left_join(site_by_mpg %>% rename(release_site = 'Release site'), 
              by = "release_site") %>%
    arrange(mpg) %>%
    ggplot(aes(x = site, y = marray_est)) + 
    geom_errorbar(aes(ymin = marray_lcl, ymax = marray_ucl, color = mpg), width = 0.1, linewidth = 1) +
    geom_point(size = 4, aes(fill = mpg), shape = 21) + 
    plot_theme + 
    facet_wrap(~fct_inorder(release_site), scales = "free_x", ncol = 3) +
    scale_y_continuous(limits = c(0, 1)) + 
    scale_color_viridis_d(option = "viridis", name = "MPG") +
    scale_fill_viridis_d(option = "viridis", name = "MPG") +
    theme(strip.background = element_rect(fill = "grey", color = "black"), 
          axis.text.x = element_text(angle = 45, hjust = 1)) + 
    ylab("Cumulative survival") + 
    xlab("To site")

```

```{r surv-plot, echo = F, fig.width=12, fig.height=20, fig.cap="Reach-specific survival estimates for juvenile sp/sum Chinook salmon during downstream migration to Lower Granite Dam. Estimates were produced using the M-array formulation of the CJS model; error bars indicate 95% confidence intervals. The color indicates the major population group."}
surv_plot
```

```{r det-plot, echo=F, fig.width=12, fig.height=18, fig.cap="Detection probability estimates for juvenile sp/sum Chinook salmon during downstream migration to Lower Granite Dam by reach. Estimates were produced using the M-array formulation of the CJS model; error bars indicate 95% confidence intervals. The color indicates the major population group."}
detection_plot
```

```{r cumulative-plot, echo=F, fig.width=12, fig.height=18, fig.cap="Estimates of cumulative survival for juvenile sp/sum Chinook salmon to Lower Granite Dam by reach. Estimates were produced using the M-array formulation of the CJS model; error bars indicate 95% confidence intervals. The color indicates the major population group."}
cumulative_plot
```

#### References
